Supermassive black holes and their host spheroids \\ I. Galaxy vivisection

### 2015-10-09 papers/data_paper/datapaper.tex
friday and i love u kri

Supermassive black holes and their host bulges/spheroids\footnote{\bf By the term ``bulge/spheroid'' 
we mean either an elliptical galaxy or the bulge component of disk galaxies, 
with no attempt of distinguishing between classical and disk-like pseudo-bulges. } 

@@

\section{Analysis}
\label{sec:anal}
\subsection{Isophotal analysis}
We performed an isophotal analysis of our galaxies using the IRAF task {\tt ellipse} \citep{taskellipse}, 
which fits elliptical isophotes to galaxy images.
The center of the isophotes was held fixed, while the ellipticity ($\epsilon$), the position angle ($P.A.$),  
and the amplitude of the fourth harmonic\footnote{The amplitude of the fourth harmonic deviations from perfect ellipses 
($B4$) parameterizes the diskyness ($B4>0$) or boxyness ($B4<0$) of the isophotes.} ($B4$) were allowed to vary with radius.
The step in semi-major axis length between successive ellipses was first set to increase linearly, 
and then geometrically in our second run\footnote{In the case of linear steps, 
the semi-major axis length for the next ellipse was calculated by adding $1$ pixel to the current length.
In the case of geometric steps, the semi-major axis length for the next ellipse 
was calculated as $1.1$ times the current length.}.
As a result, for each galaxy we produced respectively a ``linearly sampled'' and a ``logarithmically sampled'' 
surface brightness profile along the major-axis. 
Major-axis surface brightness profiles were additionally converted into the equivalent-axis,
i.e.~the geometric mean of the major ($a$) and minor ($b$) axis ($R_{\rm eq} = \sqrt{ab}$), 
equivalent to the circularized radius.
This resulted in 4 profiles per galaxy.
Isophotes corresponding to an intensity less than 
three times the root mean square of the sky background fluctuations ($3 \times RMS_{\rm sky}$) were ignored.
Some surface brightness profiles were truncated at our discretion before the $3 \times RMS_{\rm sky}$ limit, 
according to specific technical reasons (e.g.~contamination from light of a neighboring galaxy, 
disturbed morphology in the galaxy outskirts, etc.). 
{\bf In particular, we did not attempt to fit down- or up-bendings of large-scale disks 
(e.g.~\citealt{erwin2005,erwin2008,erwin2012,gutierrez2011,comeron2012,munozmateos2013,kim2014}) 
by truncating the surface brightness profiles well before the occurrence of these features. }
Individual cases are discussed in Section \ref{sec:indgal}.

@@

{\bf As stressed in Section \ref{sec:int}, our investigation is primarily focused on the central bulge/spheroidal components 

@@

unlike ``late-type spiral'' galaxies that can be bulge-less (e.g.~NGC 300).   

@@

{\bf Savorgnan \& Graham (\emph{submitted}) extensively discuss the topic of galaxies with intermediate-scale disks 

@@

(NGC 4151 is the only case for which we described the bar with a $n \sim 0.2$ S\'ersic model). 
Disk-like components embedded in the bulges of spiral galaxies were described with a $n \lesssim 1$ S\'ersic model or, 
in a few cases, with a Ferrer function\footnote{\bf The advantage of choosing a Ferrer function over a S\'ersic profile 

@@

Our galaxy sample lacks of ``late-type spiral galaxies''.
As expected, we did not find evidence of bars with exponential profiles (e.g.~\citealt{elmegreenelmegreen1985,gadotti2008,kim2015}), 
which are typical of late-type spiral galaxies \citep{gadottidesouza2006}, 
and we only modeled bars with flat profiles, which are common features of ``early-type spiral'' galaxies \citep{gadottidesouza2006}. }\\

@@

\subsection{1D versus 2D decompositions}
Here we explore how 1D and 2D decompositions compare with each other. 
Readers not interested in our practical knowledge, 
having dealt with 1D and 2D techniques of galaxy modeling at the same time, can skip to Section \ref{sec:err}. 
We summarize our experience in the following points:
\begin{itemize}
\item A visual inspection of galaxy images and their unsharp masks is often not sufficient to accurately identify all of a galaxy's components. 
Weak bars and some embedded disks can easily be missed. 
Other (inclined) embedded disks can be confused with large-scale disks.
In this regards, the 1D isophotal analysis is extremely helpful. 
Local minima/maxima or abrupt changes in the ellipticity, position angle and fourth harmonic profiles contain precious information 
about a galaxy's constituents.
\item The ellipticity and position angle of triaxial spheroids can vary with radius.
The analytic functions used by 2D decomposition codes to fit galaxy components have fixed ellipticity and position angle, 
thus they cannot account for these radial gradients. 
This problem is overcome with 1D decomposition techniques.
\item The interpretation of the residual surface brightness profile is often crucial to identify the optimal decomposition for a galaxy. 
In our experience, we found that interpreting 1D residuals was easier and more productive than 2D residuals.
\end{itemize}
Although we attempted 2D modeling for the 72 galaxies in our sample, 
more than half of the 2D decompositions were not successful or did not converge to a meaningful solution. 
This should serve as a tip to users of 2D fitting codes. 
When physically meaningful spheroid parameters are required, 
the output may not be reliable and should be inspected.
For the 31 galaxies that had successful 2D decompositions, 
we compare with their 1D parameters in Figures \ref{fig:re1d2d}, \ref{fig:n1d2d} and \ref{fig:mag1d2d}.
%The software \textsc{Imfit} returns major-axis measurements of effective radii, 
%which are plotted in Figure \ref{fig:re1d2d} versus the corresponding 1D major-axis measurements.
%2D measurements of S\'ersic indices can be approximated to 1D equivalent-axis measurements, 
%and are therefore compared to them in Figure \ref{fig:n1d2d}.
%Figure \ref{fig:mag1d2d} shows 2D versus 1D measurements of bulge magnitudes. 
The agreement between 1D and 2D effective radii and magnitudes is remarkable, 
whereas a larger amount of scatter in the S\'ersic indices can be caused by the fact that 
2D measurements do not exactly correspond to 1D equivalent-axis measurements. 
No systematic effects are observed in any of these three plots, 
which indicates that 1D and 2D techniques of galaxy modeling -- when performed on the same galaxy -- 
can give consistent results. \\
In conclusion, since we found that the best-fit parameters do not depend on the decomposition method (1D or 2D) used, 
and given that we obtained more successful 1D decompositions than 2D, 
we will base our analysis on the results from the 1D fits.

@@

\subsubsection{Uncertainties on $n_{\rm sph}$}
%\label{sec:sigman}
In Figure \ref{fig:compn}, for 58 galaxies, 
we compare the measurements of the spheroid S\'ersic index obtained by different authors 
with those obtained by us. 
For each galaxy, we computed the average value $\langle \log(n_{\rm sph}) \rangle$ of the available measurements, 
and we plot it against the scatter of the individual measurements around each spheroid's $\langle \log(n_{\rm sph}) \rangle$.
The measurements are heterogeneous, 
in the sense that they were obtained from 1D or 2D decompositions of data in different wavelengths, 
and they refer either to the major-axis, the equivalent-axis or some 2D average.
In Figure \ref{fig:compn}, the black histogram shows the distribution of the scatter around $\langle \log(n_{\rm sph}) \rangle$ 
for our measurements. 
38\% of this distribution lie within $-\sigma_1^- = -0.08~\rm dex$ and $+\sigma_1^+ = +0.06~\rm dex$, 
78\% lie within $-\sigma_2^- = -0.21~\rm dex$ and $+\sigma_2^+ = +0.17~\rm dex$,
and 92\% lie within $-\sigma_3^- = -0.25~\rm dex$ and $+\sigma_3^+ = +0.25~\rm dex$. 
We elect to use $\pm \sigma_1^\pm$, $\pm \sigma_2^\pm$ and $\pm \sigma_3^\pm$ as $1\sigma$ uncertainties 
for our measurements of $\log(n_{\rm sph})$ obtained from ``grade 1'', ``grade 2'' and ``grade 3'' fits, respectively. 
{\bf This means that, if the fit of a galaxy was classified as ``grade 1'', 
the (logarithmic) value of the spheroid S\'ersic index $\tilde{n}_{\rm sph}$ and associated uncertainties would be 
$\log(\tilde{n}_{\rm sph})^{+\sigma_1^+}_{-\sigma_1^-} = \log(\tilde{n}_{\rm sph})^{+0.06}_{-0.08}$; 
if the fit of that galaxy was classified as ``grade 2'', 
then one would have $\log(\tilde{n}_{\rm sph})^{+\sigma_2^+}_{-\sigma_2^-} = \log(\tilde{n}_{\rm sph})^{+0.17}_{-0.21}$, 
and so on.  }

@@

  \begin{figure}[h]
  \begin{center}
  \includegraphics[width=\fitfigurewidth]{images/m104_1Dfit.eps}
  \caption{M104: 
  A lenticular/spiral galaxy with a partially depleted core \citep{jardel2011}. 
  The data within the innermost $6''.1$ are excluded from the fit, 
  and the most obvious elongated feature of this galaxy is modeled with a Ferrer function.
  }
  \end{center}
  \end{figure}

>>>>

Supermassive black holes and their host spheroids\footnote{\bf By the term ``spheroid'' 
we mean either an elliptical galaxy or the bulge component of a disk galaxy, 
with no attempt at distinguishing between classical bulges and disk-like pseudo-bulges. } 

@@

\section{Analysis}
\label{sec:anal}
\subsection{Isophotal analysis}
We performed an isophotal analysis of our galaxies using the IRAF task {\tt ellipse} \citep{taskellipse}, 
which fits elliptical isophotes to galaxy images.
The center of the isophotes was held fixed, while the ellipticity ($\epsilon$), the position angle ($P.A.$),  
and the amplitude of the fourth harmonic\footnote{The amplitude of the fourth harmonic deviations from perfect ellipses 
($B4$) parameterizes the diskyness ($B4>0$) or boxyness ($B4<0$) of the isophotes.} ($B4$) were allowed to vary with radius.
The step in semi-major axis length between successive ellipses was first set to increase linearly, 
and then geometrically in our second run\footnote{In the case of linear steps, 
the semi-major axis length for the next ellipse was calculated by adding $1$ pixel to the current length.
In the case of geometric steps, the semi-major axis length for the next ellipse 
was calculated as $1.1$ times the current length.}.
As a result, for each galaxy we produced respectively a ``linearly sampled'' and a ``logarithmically sampled'' 
surface brightness profile along the major-axis. 
Major-axis surface brightness profiles were additionally converted into the equivalent-axis,
i.e.~the geometric mean of the major ($a$) and minor ($b$) axis ($R_{\rm eq} = \sqrt{ab}$), 
equivalent to the circularized radius.
This resulted in 4 profiles per galaxy.
Isophotes corresponding to an intensity less than 
three times the root mean square of the sky background fluctuations ($3 \times RMS_{\rm sky}$) were ignored.
Some surface brightness profiles were truncated at our discretion before the $3 \times RMS_{\rm sky}$ limit, 
according to specific technical reasons (e.g.~contamination from light of a neighboring galaxy, 
disturbed morphology in the galaxy outskirts, etc.). 
{\bf In particular, we did not attempt to fit bends or truncations of large-scale disks 
(e.g.~\citealt{erwin2005,erwin2008,erwin2012,gutierrez2011,comeron2012,munozmateos2013,kim2014}) 
but instead truncated the surface brightness profiles before the occurrence of such features. }
Individual cases are discussed in Section \ref{sec:indgal}.

@@

{\bf As stressed in Section \ref{sec:int}, our investigation is primarily focused on the central spheroidal components 

@@

unlike ``late-type spiral'' galaxies that can be bulgeless (e.g.~NGC 300).   

@@

{\bf \cite{savorgnangraham2015ellicular} extensively discuss the topic of galaxies with intermediate-scale disks 

@@

(NGC 4151 is the only case for which we described the bar with an $n \sim 0.2$ S\'ersic model). 
Disk-like components embedded in the bulges of spiral galaxies were described with an $n \lesssim 1$ S\'ersic model or, 
in a few cases, with a Ferrer function\footnote{\bf One advantage of choosing a Ferrer function over a S\'ersic profile 

@@

Given that our galaxy sample lacks ``late-type spiral'' galaxies, 
it is not surprising \citep{gadottidesouza2006} that we did not find bars with exponential profiles 
(e.g.~\citealt{elmegreenelmegreen1985,gadotti2008,kim2015}).
The bars in our sample were found to have rather flat inner profiles, 
as is commonly found for bars in ``early-type spiral'' galaxies \citep{gadottidesouza2006}. } \\

@@

\subsection{1D versus 2D decompositions}
Here we explore how 1D and 2D decompositions compare with each other. 
Readers not interested in our practical knowledge, 
having dealt with 1D and 2D techniques of galaxy modeling at the same time, can skip to Section \ref{sec:err}. 
We summarize our experience in the following points:
\begin{itemize}
\item A visual inspection of galaxy images and their unsharp masks is often not sufficient to accurately identify all of a galaxy's components. 
Weak bars and some embedded disks can easily be missed. 
Other (inclined) embedded disks can be confused with large-scale disks.
In this regards, the 1D isophotal analysis is extremely helpful. 
Local minima/maxima or abrupt changes in the ellipticity, position angle and fourth harmonic profiles contain precious information 
about a galaxy's constituents.
\item The ellipticity and position angle of triaxial spheroids can vary with radius.
The analytic functions used by 2D decomposition codes to fit galaxy components have fixed ellipticity and position angle, 
thus they cannot account for these radial gradients. 
This problem is overcome with 1D decomposition techniques 
{\bf because the ellipticity, and additionally the deviations from elliptical isophotes, 
are efficiently included in the equivalent-axis fit (see the Appendix of \citealt{ciambur2015})}.
\item The interpretation of the residual surface brightness profile is often crucial to identify the optimal decomposition for a galaxy. 
In our experience, we found that interpreting 1D residuals was easier and more productive than 2D residuals.
\end{itemize}
Although we attempted 2D modeling for the 72 galaxies in our sample, 
more than half of the 2D decompositions were not successful or did not converge to a meaningful solution. 
This should serve as a tip to users of 2D fitting codes. 
When physically meaningful spheroid parameters are required, 
the output may not be reliable and should be inspected.
For the 31 galaxies that had successful 2D decompositions, 
we compare with their 1D parameters in Figures \ref{fig:re1d2d}, \ref{fig:n1d2d} and \ref{fig:mag1d2d}.
%The software \textsc{Imfit} returns major-axis measurements of effective radii, 
%which are plotted in Figure \ref{fig:re1d2d} versus the corresponding 1D major-axis measurements.
%2D measurements of S\'ersic indices can be approximated to 1D equivalent-axis measurements, 
%and are therefore compared to them in Figure \ref{fig:n1d2d}.
%Figure \ref{fig:mag1d2d} shows 2D versus 1D measurements of bulge magnitudes. 
The agreement between 1D and 2D effective radii and magnitudes is remarkable, 
whereas a larger amount of scatter in the S\'ersic indices can be caused by the fact that 
2D measurements do not exactly correspond to 1D equivalent-axis measurements. 
No systematic effects are observed in any of these three plots, 
which indicates that 1D and 2D techniques of galaxy modeling -- when performed on the same galaxy -- 
can give consistent results. \\
In conclusion, since we found that the best-fit parameters do not depend on the decomposition method (1D or 2D) used, 
and given that we obtained more successful 1D decompositions than 2D, 
we will base our analysis on the results from the 1D fits.

@@

\subsubsection{Uncertainties on $n_{\rm sph}$}
%\label{sec:sigman}
In Figure \ref{fig:compn}, for 58 galaxies, 
we compare the measurements of the spheroid S\'ersic index obtained by different authors 
with those obtained by us. 
For each galaxy, we computed the average value $\langle \log(n_{\rm sph}) \rangle$ of the available measurements, 
and we plot it against the scatter of the individual measurements around each spheroid's $\langle \log(n_{\rm sph}) \rangle$.
The measurements are heterogeneous, 
in the sense that they were obtained from 1D or 2D decompositions of data in different wavelengths, 
and they refer either to the major-axis, the equivalent-axis or some 2D average.
In Figure \ref{fig:compn}, the black histogram shows the distribution of the scatter around $\langle \log(n_{\rm sph}) \rangle$ 
for our measurements. 
38\% of this distribution lie within $-\sigma_1^- = -0.08~\rm dex$ and $+\sigma_1^+ = +0.06~\rm dex$, 
78\% lie within $-\sigma_2^- = -0.21~\rm dex$ and $+\sigma_2^+ = +0.17~\rm dex$,
and 92\% lie within $-\sigma_3^- = -0.25~\rm dex$ and $+\sigma_3^+ = +0.25~\rm dex$. 
We elect to use $\pm \sigma_1^\pm$, $\pm \sigma_2^\pm$ and $\pm \sigma_3^\pm$ as $1\sigma$ uncertainties 
for our measurements of $\log(n_{\rm sph})$ obtained from ``grade 1'', ``grade 2'' and ``grade 3'' fits, respectively. 
{\bf This means that, if the fit to a galaxy was classified as ``grade 1'', 
the (logarithmic) value of the spheroid S\'ersic index $\tilde{n}_{\rm sph}$ and associated uncertainties would be 
$\log(\tilde{n}_{\rm sph})^{+\sigma_1^+}_{-\sigma_1^-} = \log(\tilde{n}_{\rm sph})^{+0.06}_{-0.08}$; 
if the fit to that galaxy was classified as ``grade 2'', 
then one would have $\log(\tilde{n}_{\rm sph})^{+\sigma_2^+}_{-\sigma_2^-} = \log(\tilde{n}_{\rm sph})^{+0.17}_{-0.21}$, 
and so on.  }

@@

  \begin{figure}[h]
  \begin{center}
  \includegraphics[width=\fitfigurewidth]{images/m104_1Dfit.eps}
  \caption{M104: 
  A lenticular/spiral galaxy with a partially depleted core \citep{jardel2011}. 
  The data within the innermost $6''.1$ are excluded from the fit, 
  and the most obvious elongated feature of this galaxy is modeled with a Ferrer function. 
  The ellipticity profile indicates that the disk is of intermediate-scale 
  {\bf (for a thorough analysis of this galaxy, see \citealt{gadottisanchezjanssen2012})}.
  }
  \end{center}
  \end{figure}
